Chord: A Static and Dynamic Program Analysis Framework for Java

### 2010-04-01 main/doc/program_representation.tex

\documentclass{article}
\usepackage{hyperlatex}

%\topmargin=-0.5in
\texonly{\oddsidemargin=0in}
\texonly{\evensidemargin=0in}
\texonly{\textwidth=5.6in}
%\textheight=8.5in

\htmlmathitalic{0}
\htmltitle{User Guide}
\htmladdress{mayur.naik@gmail.com, \today}

\providecommand\Mdom{{\bf M}}
\providecommand\Idom{{\bf I}}
\providecommand\Vdom{{\bf V}}
\providecommand\Zdom{{\bf Z}}
\providecommand\Hdom{{\bf H}}
\providecommand\Edom{{\bf E}}
\providecommand\Fdom{{\bf F}}
\providecommand\Cdom{{\bf C}}
\providecommand\Tdom{{\bf T}}
\providecommand\Pdom{{\bf P}}

\providecommand\instM{{\bf instM}}
\providecommand\statM{{\bf statM}}
\providecommand\MH{{\bf MH}}
\providecommand\MI{{\bf MI}}
\providecommand\ME{{\bf ME}}

\providecommand\cha{{\bf cha}}
\providecommand\sub{{\bf sub}}
\providecommand\HT{{\bf HT}}
\providecommand\VT{{\bf VT}}
\providecommand\EF{{\bf EF}}
\providecommand\writeE{{\bf writeE}}

\providecommand\virtIM{{\bf virtIM}}
\providecommand\statIM{{\bf statIM}}
\providecommand\specIM{{\bf specIM}}

\providecommand\MmethArg{{\bf MmethArg}}
\providecommand\MmethRet{{\bf MmethRet}}
\providecommand\IinvkArg{{\bf IinvkArg}}
\providecommand\IinvkRet{{\bf IinvkRet}}
\providecommand\MordMethArg{{\bf MordMethArg}}
\providecommand\MspcMethArg{{\bf MspcMethArg}}

\providecommand\MobjVarAsgnInst{{\bf MobjVarAsgnInst}}
\providecommand\MclsVarAsgnInst{{\bf MclsVarAsgnInst}}
\providecommand\MobjValAsgnInst{{\bf MobjValAsgnInst}}
\providecommand\MgetInstFldInst{{\bf MgetInstFldInst}}
\providecommand\MputInstFldInst{{\bf MputInstFldInst}}
\providecommand\MgetStatFldInst{{\bf MgetStatFldInst}}
\providecommand\MputStatFldInst{{\bf MputStatFldInst}}

\providecommand\syncMV{{\bf syncMV}}
\providecommand\nestedLocks{{\bf nestedLocks}}

\providecommand\VH{{\bf VH}}
\providecommand\FH{{\bf FH}}
\providecommand\HFH{{\bf HFH}}
\providecommand\IM{{\bf IM}}

\providecommand\forkC{{\bf forkC}}

\providecommand\CVC{{\bf CVC}}
\providecommand\FC{{\bf FC}}
\providecommand\CFC{{\bf CFC}}
\providecommand\CICM{{\bf CICM}}
\providecommand\CHC{{\bf CHC}}

\providecommand\clinitM{{\bf clinitM}}

\providecommand\Chord{Chord}

\providecommand\NameToTrgtMap{{\tt NameToTrgtMap}}
\providecommand\NameToTaskMap{{\tt NameToTaskMap}}
\providecommand\ConsumedNameToTasksMap{{\tt ConsumedNameToTasksMap}}
\providecommand\TaskGraph{{\tt TaskGraph}}
\providecommand\TaskMapEntries{{\tt TaskMapEntries}}
\providecommand\TrgtMapEntries{{\tt TrgtMapEntries}}
\providecommand\DlogMapEntries{{\tt DlogMapEntries}}
\providecommand\Tasks{{\tt Tasks}}
\providecommand\Trgts{{\tt Trgts}}
\providecommand\ValidTasks{{\tt ValidTasks}}
\providecommand\TrgtNames{{\tt TrgtNames}}

\title{Chord: A Static and Dynamic Program Analysis Framework for Java}
\author{Mayur Naik}
\date{\today}
\begin{document}
\maketitle
\texonly{\tableofcontents}

\texonly{\newpage}
\section{Introduction}

\Chord\ is a static and dynamic program analysis framework for Java.
It has the following key characteristics:

\begin{itemize}
\item
Stand-alone: various off-the-shelf program analyses are provided (e.g., may-alias, thread-escape, datarace, deadlock, etc.).
\item
Extensible: users can build their own program analyses on top of the provided ones.
\item
Compositional: each program analysis can be written independently and yet made to interact in complex ways with other program analyses.
\item
Efficient: results computed by each program analysis are cached for reuse by other program analyses without re-computation.
\item
Flexible: a broad range of program analyses can be expressed, including those
written imperatively in Java or declaratively in Datalog, summary-based as well as cloning-based context-sensitive analyses,
iterative refinement-based analyses, client-driven analyses, and combined static and dynamic analyses.
\end{itemize}

\Chord\ is intended for use by a broad range of users:

\begin{itemize}
\item
Software engineers: Programmers wanting to use program analyses in Chord to assist with software development and testing.
\item
Program analysis appliers: Researchers with possibly a limited program analysis background wanting to build applications on top of program analyses in Chord used as black boxes.
\item
Program analysis writers: Program analysis researchers wanting to implement and evaluate new program analysis algorithms.
\end{itemize}

\Chord\ is intended to work on a variety of platforms, including Linux, Windows, and MacOS.
It has been tested most extensively on Linux.

\Chord\ is open source software distributed under the \xlink{New BSD License}{http://www.opensource.org/licenses/bsd-license.php}.
Contributions from users are welcome and encouraged.

\texonly{\newpage}
\section{Download and Installation}

Ensure that the following software is installed on your machine:

\begin{itemize}
\item
JDK 5 or higher, e.g. from \xlink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or \xlink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
\item
a C++ compiler, e.g. \xlink{GCC}{http://gcc.gnu.org/}
\item
a Make utility, e.g. \xlink{GNU Make}{http://www.gnu.org/software/make/}
\item
\xlink{Cygwin}{http://www.cygwin.com/}, if it is Windows machine
\end{itemize}

Download the latest \Chord\ version from \xlink{here}{http://code.google.com/p/jchord/downloads/list} and uncompress it.

Directory {\tt chord/main} (henceforth called the main directory) contains a {\tt build.xml} file which is interpreted by Apache Ant.
To see the various targets available, run the following command in that directory:

\begin{quote}
\begin{verbatim}
prompt> ant
\end{verbatim}
\end{quote}

To compile \Chord, run the following command in the same directory:

\begin{quote}
\begin{verbatim}
prompt> ant compile
\end{verbatim}
\end{quote}

This will compile the following:
\begin{itemize}
\item
the Java source code of \Chord\ from {\tt chord/main/src/java/} to Java bytecode in {\tt chord/main/classes/}
\item
the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} from {\tt chord/main/src/bdd/}
to a shared library in {\tt chord/main/lib/} ({\tt libbuddy.so} on Linux, {\tt buddy.dll} on Windows, and {\tt libbuddy.dylib} on MacOS);
this library is needed for executing program analyses written in Datalog using
\xlink{bddbddd}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
the C++ source code of the Chord instrumentation agent from {\tt chord/main/src/agent/} to a shared
library in {\tt chord/main/lib} ({\tt libchord\_instr\_agent.so} on Linux, {\tt chord\_instr\_agent.dll} on Windows,
and {\tt libchord\_instr\_agent.dylib} on MacOS); this agent is needed for building program scope dynamically
and for executing dynamic program analyses.
\end{itemize}

\texonly{\newpage}
\section{A Quick Tour}

Suppose the program to be analyzed has the following directory structure:

\begin{quote}
\begin{verbatim}
example/
    src/
        foo/
            Main.java
            ...
    classes/
        foo/
            Main.class
            ...
    lib/
        src/
            taz/
                ...
        jar/
            taz.jar
    chord.properties
\end{verbatim}
\end{quote}

The above structure is typical: the program's Java source files are under \verb+example/src/+,
its class files are under \verb+example/classes/+, the source and jar files of the libraries
used by the program are under \verb+example/lib/src/+ and \verb+example/lib/jar/+, respectively.

File \verb+chord.properties+ specifies properties to be passed to Chord (alternatively,
these properties may be passed on the command-line, in format \verb+-D<name>=<value>+).
A sample such file for the above program is as follows:

\begin{quote}
\begin{verbatim}
chord.main.class=foo.Main
chord.class.path=classes:lib/jar/taz.jar
chord.src.path=src:../lib/src
chord.run.ids=0,1
chord.args.0="-thread 1 -n 10"
chord.args.1="-thread 2 -n 50"
\end{verbatim}
\end{quote}

Each relative (instead of absolute) path element in the value of any property
named \verb+chord.<...>.path+ set in this file is converted to an absolute path element with respect
to the directory containing this file, which in the above case is \verb+example/+.

Section \ref{sec-sys-props} presents an exhaustive
listing of properties recognized by \Chord.
Here, we only describe those set in the above
sample \verb+chord.properties+ file:

\begin{itemize}
\item
\verb+chord.main.class+ specifies the fully-qualified name of the main class of the
program.
\item
\verb+chord.class.path+ specifies the classpath of the program.
The value of pre-defined property \verb+${sun.boot.class.path}+ is implicitly appended
to it.
\item
\verb+chord.src.path+ specifies the Java source path of the program.
All program analyses in Chord operate on Java bytecode (specified by \verb+chord.class.path+).
The only use of this property is to HTMLize the Java source files of the program so that the
results of program analyses can be reported at the Java source code level.
\item
\verb+chord.run.ids+ specifies a list of IDs to identify runs of the program.
It is used by dynamic program analyses to determine how many times the program must be run.
An additional use of this property is to allow specifying the command-line arguments to use
in the run having ID \verb+XXX+ via property \verb+chord.args.XXX+, as illustrated by
properties \verb+chord.args.0+ and \verb+chord.args.1+ in the above example.
\end{itemize}

Execute Chord on the above program by running the following command in Chord's main directory:

\begin{quote}
\begin{verbatim}
prompt> ant -Dchord.work.dir=<...>/example run
\end{verbatim}
\end{quote}

where \verb+<...>+ denotes either the absolute path, or the path relative to Chord's main directory,
of the parent of directory {\tt example/}.
System property \verb+chord.work.dir+ specifies the working directory during Chord's execution.

The above command causes Chord to load the \verb+chord.properties+ file if present.
The location of this file may be specified explicitly by property
\verb+chord.props.file+ on the above command line.
By default, it is \verb+${chord.work.dir}/chord.properties+.

In the above case, the above command does not do much beyond loading the sample \verb+chord.properties+ file.
For Chord to do something interesting, you need to set additional properties, either
on the above command line or in the \verb+chord.properties+ file, specifying the task(s)
Chord must perform.
All tasks are summarized in the ``Chord task properties" section at \ref{sec-sys-props}.
The two most common tasks, described next, are building program scope and
running program analyses.

\paragraph{Building Program Scope}

Chord computes program scope (i.e. reachable classes and methods of a given Java program) if property
\verb+chord.build.scope+ is set to \verb+true+.
The algorithm used to compute program scope is specified by 
properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.

\begin{itemize}
\item
If property \verb+chord.reuse.scope+ is set to \verb+true+ and the files specified by
properties \verb+chord.classes.file+ and \verb+chord.methods.file+ exist,
then Chord regards those files as specifying the classes and methods, respectively,
to be regarded as reachable.  The format
of the classes file is a class name per line (e.g. \verb+foo.Main+).  The format
of the methods file is an entry of the form \verb+<method name>:<method descriptor>@<class name>+ per line,
specifying the method's name, the method's descriptor, and the method's declaring class
(e.g. \verb+main:([Ljava/lang/String;)V@foo.Main+).
Chord throws a runtime exception if the declaring class of a method in the methods file is not
specified in the classes file.
\item
If property \verb+chord.reuse.scope+ is set to \verb+false+ or the file specified by
property \verb+chord.classes.file+ or \verb+chord.methods.file+ does not exist,
then Chord computes program scope
using the algorithm specified by property \verb+chord.scope.kind+ and then
writes the classes and methods deemed reachable by that algorithm to those files.

The possible legal values of property \verb+chord.scope.kind+ are \verb+rta+ and \verb+dynamic+.
In both cases, Chord expects properties \verb+chord.main.class+ and \verb+chord.class.path+
to be set.
\begin{itemize}
\item
The \verb+rta+ value instructs Chord to compute the program scope statically using Rapid Type Analysis.
\item
The \verb+dynamic+ value instructs Chord to compute the program scope dynamically, by running the program
and observing using JVMTI the classes that are loaded at run-time.
The number of times the program is run and the command-line arguments to be supplied to
the program in each run is specified by properties \verb+chord.run.ids+ and
\verb+chord.args.XXX+ for each run ID \verb+XXX+.  By default, the program is run only once, using run ID \verb+0+,
and without any command-line arguments.
Only classes loaded in some run are regarded as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.
The rationale behind this decision is to both reduce run-time overhead of JVMTI and to increase the
predictive power of program analyses performed using the computed program scope.
\end{itemize}
\end{itemize}

The default value of \verb+chord.reuse.scope+ is \verb+false+ and the default value of \verb+chord.scope.kind+
is \verb+rta+.
Finally, the default values of \verb+chord.classes.file+ and \verb+chord.methods.file+ are
\verb+${chord.out.dir}/classes.txt+ and \verb+${chord.out.dir}/methods.txt+, respectively,
where property \verb+chord.out.dir+ defaults to \verb+${chord.work.dir}/chord_output/+,
and property \verb+chord.work.dir+ defaults to the current working directory.

\paragraph{Running Program Analyses}

@@

%chord.run.analyses
%List of names of program analyses to be run in order.
%chord.java.analysis.path	 path	 Classpath containing program analyses in Java (i.e. @Chord-annotated classes).	${chord.main.dir}/classes/
%chord.dlog.analysis.path	 path	 Path of directories containing program analyses in Datalog (i.e. *.datalog and *.dlog files).	${chord.main.dir}/src/main/dlog/
%chord.reuse.rels	 bool	 Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them.	false
%chord.publish.results	 bool	 Publish the results of program analyses in HTML.	true

%\texonly{\newpage}
%\section{Program Representation (todo)}

%TODO

%\texonly{\newpage}
%\section{Implementing Program Analyses (todo)}

%TODO

\texonly{\newpage}
\section{Chord System Properties}
\label{sec-sys-props}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.

\begin{table}[htp]
%\caption{}
\xmlattributes*{table}{border="1"}
\begin{center}
\begin{tabular}{|l|lll|}
\hline
{\bf Property name} & {\bf Type} & {\bf Description} & {\bf Default value (if any)} \\
\hline
\multicolumn{4}{l}{{\bf Chord resource properties}} \\
\verb+chord.main.dir+ & location & \verb+main+ directory in Chord's installation. & \\
\hline
\verb+chord.lib.dir+ & location & Directory containing libraries needed by Chord. & \verb+${chord.main.dir}/lib/+ \\
\hline
\verb+chord.main.class.path+ & path & Classpath of Chord. It includes the path specified by property \verb+chord.java.analysis.path+ to allow running user-defined program analyses. & See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.bddbddb.class.path+ & path & Classpath of bddbddb. & See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.instr.agent.file+ & location & Shared library implementing Chord instrumentation agent. & \verb+${chord.main.dir}/lib/libchord_instr_agent.so+ (on Linux; \verb+.dll+ on Windows and \verb+.dylib+ on MacOS)\\
\hline
\verb+chord.javadoc.url+ & string & URL of the Javadoc location of program analyses. It is used when publishing targets (i.e. when property \verb+chord.publish.targets+ is set to true). & \verb+http://chord.stanford.edu/javadoc/+ \\
\hline
\multicolumn{4}{l}{{\bf Chord boot properties}} \\
\hline
\verb+chord.work.dir+ & location & Working directory during Chord's execution. & current working directory \\
\hline
\verb+chord.props.file+ & location & Properties file loaded by \verb+${chord.main.dir}/build.xml+.  Any of the below properties, as well as other user-defined properties to be passed to Chord (e.g. for user-defined analyses), may be set in this file.  Each relative (instead of absolute) path element in the value of any property named \verb+chord.<...>.path+ set in this file is converted to an absolute path element with respect to the directory containing this file. & \verb+${chord.work.dir}/chord.properties+ \\
\hline
\verb+chord.max.heap+ & string & Maximum memory size of JVM running Chord. & \verb+1024m+ \\
\hline 
\verb+chord.max.stack+ & string & Maximum thread stack size of JVM running Chord. & \verb+32m+ \\
\hline
\verb+chord.jvmargs+ & string & Arguments passed to JVM running Chord. & \verb+"-showversion -ea -Xmx${chord.max.heap} -Xss${chord.max.stack}"+ \\
\hline
\verb+chord.bddbddb.max.heap+ & string & Maximum memory size of JVM running bddbddb. & \verb+1024m+ \\
\hline
\multicolumn{4}{l}{{\bf Chord task properties}} \\
\hline
\verb+chord.build.scope+ & bool & Compute program scope (i.e. reachable classes and methods) using algorithm specified by properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.  Program scope is computed regardless of the value of this property if another task (e.g. a program analysis to be run) needs it. & \verb+false+ \\
\hline
\verb+chord.run.analyses+ & string list & List of names of program analyses to be run in order. & \verb+""+ \\
\hline
\verb+chord.print.rels+ & string list & List of names of program relations whose contents must be printed to files \verb+${chord.out.dir}/<...>.txt+ where \verb+<...>+ denotes the relation name. & \verb+""+ \\
\hline
\verb+chord.publish.targets+ & bool & Publish all targets defined by program analyses in paths specified by properties \verb+chord.java.analysis.path+ and \verb+chord.dlog.analysis.path+. & \verb+false+ \\
\hline
\multicolumn{4}{l}{{\bf Basic program properties}} \\
\hline
\verb+chord.main.class+ & class & Fully-qualified name of main class of given program (e.g. \verb+com.example.Main+). &  \\
\hline
\verb+chord.class.path+ & path & Classpath of given program. \verb+${sun.boot.class.path}+ is implicitly appended to it. & \verb+${sun.boot.class.path}+ \\
\hline
\verb+chord.src.path+ & path & Java source path of given program. & \verb+""+ \\
\hline
\verb+chord.run.ids+ & string list & List of IDs to identify runs of given program. & \verb+0+ \\
\hline
\verb+chord.args.XXX+ & string & Command-line argument string to be used for given program in run having ID \verb+XXX+. & \verb+""+ \\
\hline
\verb+chord.runtime.jvmargs+ & string & Arguments to JVM when running given program.  It is used when the program is executed to compute its scope (i.e. when property \verb+chord.scope.kind+ is set to \verb+dynamic+) as well as when the instrumented program is executed for dynamic analyses. & \verb+"-ea -Xmx1024m"+ \\ 
\hline
\multicolumn{4}{l}{{\bf Program scope properties}} \\ 
\hline
\verb+chord.scope.kind+ & [\verb+rta+|\verb+dynamic+] & Algorithm to compute program scope (i.e. reachable classes and methods).  Current choices include \verb+rta+ (static Rapid Type Analysis) and \verb+dynamic+ (dynamic analysis). & \verb+rta+ \\
\hline
\verb+chord.reuse.scope+ & bool & Treat program scope as the classes and methods listed in files specified by properties \verb+chord.classes.file+ and \verb+chord.methods.file+, respectively. Property \verb+chord.scope.kind+ is ignored if this property is set to \verb+true+ and the two files exist. & \verb+false+ \\
\hline
\verb+chord.check.exclude+ & string list & List of prefixes of names of classes and packages to be excluded by program analyses. Interpretation of this property is analysis-specific. & \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\verb+chord.instr.exclude+ & string list & List of prefixes of names of classes and packages to be excluded from being instrumented.  Classes in package \verb+java.lang.ref+ are not instrumented regardless of the value of this property. & \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\multicolumn{4}{l}{{\bf Program analysis properties}} \\
\hline
\verb+chord.java.analysis.path+ & path & Classpath containing program analyses written in Java (i.e. \verb+@Chord+-annotated classes). & \verb+${chord.main.dir}/classes/+ \\
\hline
\verb+chord.dlog.analysis.path+ & path & Path of directories containing program analyses written in Datalog (i.e. \verb+*.datalog+ and \verb+*.dlog+ files). & \verb+${chord.main.dir}/src/dlog/+ \\
\hline
\verb+chord.reuse.rels+ & bool & Construct program relations from BDDs stored on disk (from a previous run of Chord) whenever possible instead of re-computing them. & \verb+false+ \\
\hline
\verb+chord.publish.results+ & bool & Publish the results of program analyses in HTML. Interpretation of this property is analysis-specific. & \verb+true+ \\
\hline
\multicolumn{4}{l}{{\bf Program transformation properties}} \\
\hline
\verb+chord.ssa+ & bool & Do SSA transformation for all methods deemed reachable by the algorithm used to compute program scope. & \verb+true+ \\
\hline
\multicolumn{4}{l}{{\bf Chord debug properties}} \\
\hline
\verb+chord.bddbddb.noisy+ & bool & Make bddbddb output debug info when solving Datalog programs. & \verb+false+ \\
\hline
\verb+chord.save.maps+ & bool & Write to file \verb+${chord.bddbddb.work.dir}/<...>.map+ when saving program domain named \verb+<...>+ (useful for debugging Datalog programs using \verb+debug+ target of \verb+${chord.main.dir}/build.xml+). & \verb+true+ \\
\hline
\multicolumn{4}{l}{{\bf Chord instrumentation properties}} \\
\hline
\verb+chord.trace.pipe+ & bool & Implement the dynamic trace file as a POSIX pipe instead of a regular file. & \verb+true+ \\
\hline
\verb+chord.trace.block.size+ & int & Number of bytes to read/write in a single operation from/to the dynamic trace file. & \verb+4096+ \\
\hline
\verb+chord.runtime.class+ & class & Subclass of \verb+chord.project.Runtime+ used by dynamic program analyses to handle events generated during an instrumented program's execution. & \verb+chord.project.BufferedRuntime+ \\
\hline
\hline
\multicolumn{4}{l}{{\bf Chord output properties}} \\
\hline
\verb+chord.out.dir+ & location & Directory to which Chord dumps all files. & \verb+${chord.work.dir}/chord_output/+ \\
\hline
\verb+chord.out.file+ & location & File to redirect standard output stream during Chord's execution. & \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.err.file+ & location & File to redirect standard error stream during Chord's execution. & \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.classes.file+ & location & File from/to which list of classes deemed reachable is read/written. & \verb+${chord.out.dir}/classes.txt+ \\
\hline
\verb+chord.methods.file+ & location & File from/to which list of methods deemed reachable is read/written. & \verb+${chord.out.dir}/methods.txt+ \\
\hline
\verb+chord.bddbddb.work.dir+ & location & Directory used by bddbddb as its input/output directory. & \verb+${chord.out.dir}/bddbddb/+ \\
\hline
\verb+chord.boot.classes.dir+ & location & Directory containing instrumented JDK classes used by given program. & \verb+${chord.out.dir}/boot_classes/+ \\
\hline
\verb+chord.user.classes.dir+ & location & Directory containing instrumented non-JDK classes of given program. & \verb+${chord.out.dir}/user_classes/+  \\
\hline
\verb+chord.instr.scheme.file+ & location & File specifying kind and format of events in dynamic trace file. & \verb+${chord.out.dir}/scheme.ser+ \\
\hline
\verb+chord.crude.trace.file+ & location & Crude dynamic trace file. & \verb+${chord.out.dir}/crude_trace.txt+ \\
\hline
\verb+chord.final.trace.file+ & location & Final dynamic trace file. & \verb+${chord.out.dir}/final_trace.txt+
\end{tabular}
\end{center}
\end{table}

\texonly{\newpage}
\section{Acknowledgments}

Chord would not be possible without the following open-source software:

\begin{itemize}
\item
\xlink{Joeq}{http://joeq.sourceforge.net/}, a Java compiler framework
\item
\xlink{Javassist}{http://www.csg.is.titech.ac.jp/~chiba/javassist/}, a Java bytecode manipulation tool
\item
\xlink{bddbddb}{http://bddbddb.sourceforge.net/}, a BDD-based Datalog solver
\end{itemize}

Chord additionally uses the following open-source tools and libraries:
\begin{itemize}
\item
\xlink{Ant-Contrib}{http://ant-contrib.sourceforge.net/}, a collection of useful Ant tasks  
\item
\xlink{BuDDy}{http://buddy.sourceforge.net/}, a BDD library
\item
\xlink{GNU Trove}{http://trove4j.sourceforge.net/}, a primitive collections library for Java
\item
\xlink{Java2HTML}{http://www.java2html.com/} and \xlink{Java2Html}{http://www.java2html.de/}, Java to HTML tools
\item
\xlink{Saxon}{http://saxon.sourceforge.net/}, an XSLT processor
\end{itemize}

Chord was supported in part by grants from the National Science Foundation, an equipment grant from Intel, and a Microsoft fellowship during 2005-2007.

\end{document}

\texonly{\newpage}
\section{Program Representation}
\label{sec-prog-build}

Given the program's application classpath via system property {\tt chord.app.class.class}
and either the program's main class via system property {\tt chord.main.class} or
the program's root methods via system property {\tt chord.root.classes}
(and, optionally, system properties {\tt chord.include.root.methods}
and {\tt chord.exclude.root.methods}),
\Chord\ uses system property {\tt chord.program.builder.class} to locate the
builder to be used for building an intermediate representation of the program,
namely, an object of class {\tt chord.program.Program}.
By default, the value of this property is class {\tt chord.program.builders.SootProgramBuilder}
which uses the Soot Java compiler framework \cite{Soot}.
Any builder, however, must implement interface
{\tt chord.program.builders.IProgramBuilder}.

The intermediate representation
elides values of primitive type (e.g., int, boolean, etc.) and operations
on them (e.g., arithmetic and relational operations) in the original program.
The current emphasis of \Chord\ is on the analysis of objects
but the analysis of primitives may be supported in future
releases.

The intermediate representation represents the program as a set of types.
Each type has a set of (instance or static) fields and a set of
(instance or static) methods.
Each method has an optional control-flow graph.
The control-flow graph has a unique entry point and
a unique exit point\footnote{Even the control-flow graph of a method with an infinite loop
has an exit point because loops are transformed into tail-recursive methods
(see Section \ref{sec-prog-transform}).}.
The entry and exit points are always distinct.
Each point in the control-flow graph has one or more immediate predecessors
(except the entry point which has none) and, likewise,
one or more immediate successors (except the exit point which has none).
Associated with each point is a simple statement, which may be of one of
the following kinds:

\begin{itemize}
\item
{\tt HeadInst}

The statement associated with the unique entry point of a control-flow graph.

It serves as a placeholder for:

\begin{itemize}
\item
the (possibly empty) list of the method's {\it argument variables} of reference type;
in the case of an instance method, this list is non-empty,
with the $0^{th}$ element being the distinguished implicit {\tt this}
argument variable.
\item
the (possibly empty) list of the method's {\it temporary variables} of reference type.
\end{itemize}

Argument variables and temporary variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

Argument variables and temporary variables are distinct, namely,
the set of a method's {\it local variables} (of reference type)
is the disjoint union of the sets of the method's argument variables (of reference type) and
temporary variables (of reference type).

\item
{\tt TailInst}

The statement associated with the unique exit point of a control-flow graph.

It serves as a placeholder for 
the (possibly empty) list of the method's {\it return variables} of reference type.

Return variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

The set of a method's return variables (of reference type) is a subset of the set
of the method's local variables (of reference type).

A Java method has at most one return variable but
\Chord's intermediate representation allows multiple return variables
to enable compiling away exceptions, loops, synchronized blocks, etc.

\item
{\tt AcqLockInst}

A statement of the form {\tt monitorenter v} where
{\tt v} is a local variable of reference type.

\item
{\tt RelLockInst}

A statement of the form {\tt monitorexit v} where {\tt v}
is a local variable of reference type.

\item
{\tt StrValAsgnInst}

A statement of the form {\tt v = "..."} where {\tt v} is
a local variable of any reference type that can store a string
constant.  It is typically (but not necessarily) of type
{\tt java.lang.String}.

\item
{\tt NilValAsgnInst}

A statement of the form {\tt v = null} where {\tt v}
is a local variable of reference type.

\item
{\tt StatFldRefInst}

A statement of one of the following forms:

{\tt v = f} \\
{\tt f = v} \\
{\tt * = f} \\
{\tt f = *}

where {\tt v} is a local variable of reference type and
{\tt f} is a static field.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.

\item
{\tt InstFldRefInst}

A statement of one of the following forms:

{\tt v = b.f} \\
{\tt b.f = v} \\
{\tt * = b.f} \\
{\tt b.f = *}

where {\tt v} and {\tt b} are local variables of reference
type and {\tt f} is an instance field.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.

\item
{\tt SkipInst}

A no-op statement.

Due to \Chord's emphasis on the analysis of objects, each statement
in the original program that does not access
any value of reference type (i.e., only accesses values of primitive
type) is represented by this statement.

\item
{\tt InvkInst}

A method invocation statement of the form
{\tt r1,...,rn = m(a1,...,ak)} where
{\tt r1}, ..., {\tt rn} (return variables) and
{\tt a1}, ..., {\tt ak} (argument variables)
are local variables of reference type.

Argument variables and return variables of primitive type are
elided due to \Chord's emphasis on the analysis of objects.

A method invocation statement in Java has at most
one return variable but
\Chord's intermediate representation allows multiple return variables
to enable compiling away exceptions, loops, synchronized blocks, etc.

\item
{\tt ObjVarAsgnInst}

A statement of the form {\tt v1 = v2} where {\tt v1}
and {\tt v2} are local variables of (possibly different) reference types.

\item
{\tt ObjValAsgnInst}

A statement of the form {\tt v = new t} where {\tt t} is a
concrete reference type and
{\tt v} is a local variable of a (possibly different) reference type.

\item
{\tt ClsVarAsgnInst}

A statement of the form {\tt v = b.getClass()} where
{\tt v} is a local variable of type {\tt java.lang.Class}
and {\tt b} is a local variable of reference type.

The reason this kind of statement is modeled is because
\Chord\ has an emphasis on concurrency and many Java programs
hold locks on the runtime class of an object
(which is obtained by calling the native {\tt getClass}
instance method of the object).

\item
{\tt AryElemRefInst}

A statement of one of the following forms:

{\tt v = b[*]}  \\
{\tt b[*] = v} \\
{\tt * = b[*]} \\
{\tt b[*] = *}

where {\tt v} and {\tt b} are local variables of reference
type.

The {\tt *} denotes an expression of primitive type in the
original program which is elided due to \Chord's emphasis on the
analysis of objects.
\end{itemize}

\texonly{\newpage}
\section{Task Graph Construction}
\label{sec-tg}

\Chord\ requires the following three system properties to construct the task graph:

\begin{enumerate}
\item
{\tt chord.dlogmap.path}

The path specifying the mapping from the name of each Datalog-based program
analysis that must be included in the project to the location
of the file defining that program analysis.

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <file-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:' on UNIX and `;' on Windows).

Let \DlogMapEntries\ denote the set of all entries of the
above form.

\item
{\tt chord.trgtmap.path}

The path partially specifying the mapping from the name of each program domain,
program relation, and Java-based program analysis that must be
included in the project to the fully-qualified name of a subclass, either
pre-defined in \Chord\ or defined by the user, of {\tt chord.Dom},
{\tt chord.Ary*Rel}, and
{\tt chord.Entity}, respectively, the instance fields of which define the
state of the corresponding domain, relation, or analysis.
The `{\tt *}' in {\tt chord.Ary*Rel} is 1, 2, ..., 6, or N for
program relations of arity 1, 2, ..., 6, or $>$ 6, respectively.

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:` on UNIX and ';' on Windows).

Let \TaskMapEntries\ denote the set of all entries of the
above form.

\item
{\tt chord.taskmap.path}

The path partially specifying the mapping from the name of each program domain,
program relation, and Java-based program analysis that must be included in
the project to the fully-qualified name of a
subclass, either pre-defined in \Chord\ or defined by the user, of
{\tt chord.Dom}, {\tt chord.Ary*Rel}, and {\tt chord.Entity}, respectively,
which define the following instance methods:
\begin{itemize}
\item
the {\tt execute} method specifying
the computation that updates the state of the corresponding domain, relation,
or analysis,
\item
the {\tt invalidate} method specifying the action to be taken when the
state of the corresponding domain, relation, or analysis is invalidated.
\item
the {\tt getConsumedNames} method specifying names of any domains, relations,
and Datalog-based or Java-based program analyses upon whose state the
computation depends, and
\item
the {\tt getProducedNames} method specifying names of any domains, relations,
and Datalog-based or Java-based program analyses whose state depends upon
the computation.
\end{itemize}

Each path element must be the location of a file containing a
list of entries (one per line) of the form:

{\tt <name> <class-name>}

The path elements must be separated by the system-specific path separator
character (e.g., `:` on UNIX and ';' on Windows).

Let \TrgtMapEntries\ denote the set of all entries of the above form.
\end{enumerate}

Then, \Chord\ uses the values of \DlogMapEntries, \TaskMapEntries, and \TrgtMapEntries\ 
to do the following:

\begin{enumerate}
\item
Compute {\tt Tasks\ :\ Set<chord.Entity>}, the set of all tasks,
each of which is an object of (a subclass of) {\tt chord.Entity},
as follows:

\begin{enumerate}
\item
For each entry in \DlogMapEntries\ (i.e., a Datalog-based program analysis,
there is a unique task, namely, a unique object of class {\tt chord.DatalogAnalysis}.

\item
For each entry in \TaskMapEntries\ (i.e., of the form {\tt <name> <class-name>} specifying
a program domain, a program relation, or a Java-based program analysis), there is a
unique task, namely, a unique object
of the class named {\tt <class-name>} which must be (a subclass of) {\tt chord.Dom} in the
case of a program domain, (a subclass of) {\tt chord.Ary*Rel} in the case of a
program relation, and (a subclass of) {\tt chord.Entity} in the case of a Java-based
program analysis.
\end{enumerate}

\item
Compute the following maps as shown below:

{\tt ConsumedNameToTasksMap\ :\ String -> Set<Task>},
mapping each name to the set of (possibly empty) tasks that consume that name.

{\tt NameToTaskMap\ :\ String -> Task},
mapping each name to the unique task that produces that name.

>>>>

\section{Program Representation}
\label{sec:program-representation}

@@
